---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# library(readr)
library(tidyverse)
library(stringr)
library(MPRAnalyze)
library(BiocParallel)
library(pheatmap)
library(GGally)
library(here)
```


Read in counts file
```{r}
lib <- read_tsv(here("data/nextseq_counts_111920.tsv"))
```

Separate oligo name into locus, allele, barcode
```{r}
lib <- lib %>% separate(oligo, into=c("oligo_type", "locus", "allele", "barcode"), 
                        sep="_", remove=F) %>%
  mutate(locus = paste(oligo_type, locus, sep="_")) %>%
  dplyr::select(-oligo, -oligo_type)
```


```{r}
lib_grp <- lib %>% group_by(locus, allele) %>% 
  summarise(num_plasmid_nonzero = sum(plasmid_2 > 0 | plasmid_3 > 0))
lib_grp_filt <- dplyr::filter(lib_grp, num_plasmid_nonzero < 3)
lib_filt <- dplyr::filter(lib, !(locus %in% lib_grp_filt$locus))

```

```{r}

# take out scrambles and motifs, then add reformatted scrambles back in
# filter out the non-snps
lib_filt <- dplyr::filter(lib_filt, !grepl("JM|motif|scramble", lib_filt$locus))
```

Make DNA matrix for MPRAnalyze
```{r}
dna <- lib_filt %>% dplyr::select(locus, allele, barcode, plasmid_2, plasmid_3)%>%
  pivot_longer(cols=c(plasmid_2, plasmid_3), names_to = "dna_batch", values_to = "count") %>%
  mutate(dna_batch = ifelse(dna_batch == "plasmid_2", "rep1", "rep2")) %>% rowwise() %>%
  mutate(batch.allele.bc = paste(dna_batch, allele, barcode, sep="-")) %>%
  dplyr::select(-barcode, -allele, -dna_batch) %>%
  spread(key= batch.allele.bc, value = count)
future_rownames <- dna$locus
dna <- dna %>% ungroup() %>% dplyr::select(-locus)
dna[is.na(dna)] <- 0
dna <- as.matrix(dna)
rownames(dna) <- future_rownames

```

Make RNA matrix with all celltypes
```{r}
rna <- lib_filt %>% dplyr::select(-plasmid_2, -plasmid_3) %>%
  pivot_longer(c(-allele, -locus, -barcode), names_to = "cell_type",
                                                        values_to="count")
rna <- rna %>% unite("cell_type", cell_type, barcode, sep="-") %>%
  unite("cell_type", cell_type, allele, sep="-")
rna <- rna %>% pivot_wider(names_from=cell_type, values_from = count)

future_rownames <- rna$locus
rna <- rna %>% ungroup() %>% dplyr::select(-locus)
rna[is.na(rna)] <- 0
rna <- as.matrix(rna)
rownames(rna) <- future_rownames
rna <- rna[order(as.character(rownames(rna))),]

```

Make DNA column annotations
```{r}
col_dna = data.frame(dna_batch = str_remove(colnames(dna), "-.*$"),
                     condition = str_remove(str_remove(colnames(dna), "^rep[0-9]-"), "-.*$"),
                      barcode = str_remove(colnames(dna), "^.*-"))
rownames(col_dna) <- colnames(dna)
col_dna$barcode_allelic <- interaction(col_dna$barcode, col_dna$condition)
```

I've been running MPRAnalyze separately for each cell type, made it easier to interpret, need to think more about what this means for multiple hypothesis correction
- considering adding a filter for 3 or more barcodes represented in each cell type, would need to be in the for loop
```{r}
cells = unique(str_remove(colnames(rna), "_.*"))
loci = distinct(dplyr::select(lib_filt, locus))
library(foreach)
library(doParallel)
numCores = 6
# cells = cells[c(1:6)]
registerDoParallel(numCores)  # use multicore, set to the number of our cores
res1 <- foreach (cell=cells, .combine=cbind) %dopar% {
  print(cell)
  # dna and col_dna can stay the same
  # rna needs to be filtered for only the cell related columns
  # rna col annot needs to be made
  rna_cell <- rna[,grepl(cell, colnames(rna))]
  
  col_rna = data.frame(condition = stringr::str_remove(colnames(rna_cell), "^.*-"), 
                       batch = stringr::str_remove(stringr::str_remove(colnames(rna_cell), "^.*_"), "-.*$"),
                       barcode = stringr::str_remove(stringr::str_remove(colnames(rna_cell), "-ref$|-alt$"), ".*-"))
  rownames(col_rna) <- colnames(rna_cell)
  col_rna$barcode_allelic <- interaction(col_rna$barcode, col_rna$condition)
  obj <- MPRAnalyze::MpraObject(dnaCounts =dna, rnaCounts =rna_cell, dnaAnnot = col_dna, 
                                rnaAnnot = col_rna)
  obj <- MPRAnalyze::estimateDepthFactors(obj, lib.factor = "dna_batch", which.lib = "dna", 
                                          depth.estimator = "uq")
  obj <- MPRAnalyze::estimateDepthFactors(obj,which.lib = "rna", lib.factor= "batch",
                                          depth.estimator = "uq")
  obj <- MPRAnalyze::analyzeComparative(obj = obj, 
                                       dnaDesign = ~ barcode_allelic, 
                                       rnaDesign = ~ condition, 
                                       reducedDesign = ~ 1)
# should I have batch in the rnaDesign? Try both:
#   obj <- analyzeComparative(obj = obj, 
#                           dnaDesign = ~ barcode_allelic, 
#                           rnaDesign = ~ condition + batch, 
#                           reducedDesign = ~ batch)
  res_cell <- MPRAnalyze::testLrt(obj) 
  res_cell <- dplyr::select(res_cell, fdr, pval, logFC)
  colnames(res_cell) <- paste(cell, colnames(res_cell), sep="_")
  res_cell <- dplyr::left_join(loci, tibble::rownames_to_column(res_cell, "locus"))
}

res_locus <- res1$locus
res1_write <- res1 %>% dplyr::select(-locus)
res1_write$locus <- res_locus
res1_write <- res1_write %>% dplyr::select(locus, dplyr::everything())
write_tsv(res1_write, here("output/mpranalyze_plasmidrep_112720.tsv"))

```
